Canada’s Battle Against Car Theft: Progress Amid Persistent Challenges 

### 2024-12-04 paper/paper.qmd
sketchupdate

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(here)
library(lubridate)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(distributions3)
library(topmodels)
library(MASS)
library(sf)
library(rstudioapi)
library(opendatatoronto)
library(janitor)
```

@@

variable_description <- data.frame(
  Column = c(
    "X_id", "EVENT_UNIQUE_ID", "REPORT_DATE", "OCC_DATE", "REPORT_YEAR",
    "REPORT_MONTH", "REPORT_DAY", "REPORT_DOY", "REPORT_DOW", "REPORT_HOUR",
    "OCC_YEAR", "OCC_MONTH", "OCC_DAY", "OCC_DOY", "OCC_DOW", "OCC_HOUR",
    "DIVISION", "LOCATION_TYPE", "PREMISES_TYPE", "UCR_CODE", "UCR_EXT",
    "OFFENCE", "MCI_CATEGORY", "HOOD_158", "LONG_WGS84", "LAT_WGS84"
  ),
  Description = c(
    "Unique row identifier for Open Data database",
    "Offence Number",
    "Date Offence was Reported",
    "Date of Offence",
    "Year Offence was Reported",
    "Month Offence was Reported",
    "Day of the Month Offence was Reported",
    "Day of the Year Offence was Reported",
    "Day of the Week Offence was Reported",
    "Hour Offence was Reported",
    "Year Offence Occurred",
    "Month Offence Occurred",
    "Day of the Month Offence Occurred",
    "Day of the Year Offence Occurred",
    "Day of the Week Offence Occurred",
    "Hour Offence Occurred",
    "Police Division where Offence Occurred",
    "Location Type of Offence",
    "premise types of Offence",
    "UCR Code for Offence",
    "UCR Extension for Offence",
    "Title of Offence",
    "MCI Category of Occurrence",
    "Identifier of Neighbourhood using City of Toronto's new 158 neighbourhood structure",
    "Longitude coordinate",
    "Latitude coordinate"
  )
)
# Render the table
kable(variable_description, col.names = c("Column", "Description")) %>% kable_styling(full_width = FALSE) %>%  column_spec(1, width = "4cm") %>% column_spec(2, width = "8cm")
```

@@

i_am("README.md")
dt <- read_csv(here("data","analysis_data","analysis_data.csv"))
data_part1 <- kable(head(dt)[, 1:6],caption = "Sample of Clean Data: Part 1")
data_part2 <- kable(head(dt)[, 7:10],caption = "Sample of Clean Data: Part 2")
data_part3 <- kable(head(dt)[, 11:13],caption = "Sample of Clean Data: Part 3")
data_part1
data_part2
data_part3
```

@@

monthly_counts_all <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))),
    action = ifelse(time >= as.numeric(as.Date("2024-01-01")),1,0))
poisson_model_all <- glm(
  Crime_Count ~ time +action,
  family = poisson(link = "log"),
  data = monthly_counts_all
)
kable(summary(poisson_model_all)$coefficients)
```

@@

dispersion_stat <- sum(residuals(poisson_model_all, type = "pearson")^2) / poisson_model_all$df.residual
kable(data.frame(Stats="Dispersion Factor",Value=dispersion_stat))
```

@@

kable(summary(poisson_model_all2)$coefficients)
```

@@

t1 <- as.data.frame(confint(poisson_model_all))
t2 <- as.data.frame(confint(poisson_model_all2))
t1$model_type <- "Poisson Model"
t2$model_type <- "Quasi-Poisson Model"
kable(rbind(t1,t2))
```

@@

```{r}
#| label: tbl-Q-P-PREMISES_TYPE
#| echo: false
#| warning: false
#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of premise types"
monthly_counts_PREMISES_TYPE <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH,PREMISES_TYPE) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.Date(paste0(YEARMONTH, "-01")),
    action = ifelse(time >= as.Date("2024-01-01"),1,0))
poisson_model_PREMISES_TYPE <- glm(
  Crime_Count ~ time+PREMISES_TYPE +action,
  family = quasipoisson(),
  data = monthly_counts_PREMISES_TYPE
)
kable(summary(poisson_model_PREMISES_TYPE)$coefficients)
```

```{r}
#| label: tbl-Q-P-OFFENCE
#| echo: false
#| warning: false
#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of offence types"
monthly_counts_OFFENCE <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH,OFFENCE) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.Date(paste0(YEARMONTH, "-01")),
    action = ifelse(time >= as.Date("2024-01-01"),1,0))
poisson_model_OFFENCE <- glm(
  Crime_Count ~ time+OFFENCE +action,
  family = quasipoisson(),
  data = monthly_counts_OFFENCE
)
kable(summary(poisson_model_OFFENCE)$coefficients)
```

```{r}
#| label: tbl-Q-P-DIVISION
#| echo: false
#| warning: false
#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of divisions"
monthly_counts_DIVISION <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH,DIVISION) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.Date(paste0(YEARMONTH, "-01"))  ,
    action = ifelse(time >= as.Date("2024-01-01"),1,0))
poisson_model_DIVISION <- glm(
  Crime_Count ~ time+DIVISION +action,
  family = quasipoisson(),
  data = monthly_counts_DIVISION
)
kable(summary(poisson_model_DIVISION)$coefficients)
```

>>>>

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(lubridate)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(distributions3)
library(topmodels)
library(MASS)
library(sf)
library(rstudioapi)
library(opendatatoronto)
library(janitor)
```

@@

variable_description <- data.frame(
  Column = c(
    "X_id", "EVENT_UNIQUE_ID", "REPORT_DATE", "OCC_DATE", "REPORT_YEAR",
    "REPORT_MONTH", "REPORT_DAY", "REPORT_DOY", "REPORT_DOW", "REPORT_HOUR",
    "OCC_YEAR", "OCC_MONTH", "OCC_DAY", "OCC_DOY", "OCC_DOW", "OCC_HOUR",
    "DIVISION", "LOCATION_TYPE", "PREMISES_TYPE", "UCR_CODE", "UCR_EXT",
    "OFFENCE", "MCI_CATEGORY", "HOOD_158", "LONG_WGS84", "LAT_WGS84"
  ),
  Description = c(
    "Unique row identifier for Open Data database",
    "Offence Number",
    "Date Offence was Reported",
    "Date of Offence",
    "Year Offence was Reported",
    "Month Offence was Reported",
    "Day of the Month Offence was Reported",
    "Day of the Year Offence was Reported",
    "Day of the Week Offence was Reported",
    "Hour Offence was Reported",
    "Year Offence Occurred",
    "Month Offence Occurred",
    "Day of the Month Offence Occurred",
    "Day of the Year Offence Occurred",
    "Day of the Week Offence Occurred",
    "Hour Offence Occurred",
    "Police Division where Offence Occurred",
    "Location Type of Offence",
    "premise types of Offence",
    "UCR Code for Offence",
    "UCR Extension for Offence",
    "Title of Offence",
    "MCI Category of Occurrence",
    "Identifier of Neighbourhood using City of Toronto's new 158 neighbourhood structure",
    "Longitude coordinate",
    "Latitude coordinate"
  )
)
# Render the table
kable(variable_description, col.names = c("Column", "Description"), format = "latex") %>% kable_styling(full_width = FALSE) %>%  column_spec(1, width = "4cm") %>% column_spec(2, width = "8cm")
```

@@

i_am("README.md")
dt <- read_csv(here("data","analysis_data","analysis_data.csv"))
data_part1 <- kable(head(dt)[, 1:6],caption = "Sample of Clean Data: Part 1", format = "latex")
data_part2 <- kable(head(dt)[, 7:10],caption = "Sample of Clean Data: Part 2", format = "latex")
data_part3 <- kable(head(dt)[, 11:13],caption = "Sample of Clean Data: Part 3", format = "latex")
data_part1
data_part2
data_part3
```

@@

monthly_counts_all <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))),
    action = ifelse(time >= as.numeric(as.Date("2024-01-01")),1,0))
poisson_model_all <- glm(
  Crime_Count ~ time +action,
  family = poisson(link = "log"),
  data = monthly_counts_all
)
kable(summary(poisson_model_all)$coefficients, format = "latex")
```

@@

dispersion_stat <- sum(residuals(poisson_model_all, type = "pearson")^2) / poisson_model_all$df.residual
kable(data.frame(Stats="Dispersion Factor",Value=dispersion_stat), format = "latex")
```

@@

kable(summary(poisson_model_all2)$coefficients, format = "latex")
```

@@

t1 <- as.data.frame(confint(poisson_model_all))
t2 <- as.data.frame(confint(poisson_model_all2))
t1$model_type <- "Poisson Model"
t2$model_type <- "Quasi-Poisson Model"
kable(rbind(t1,t2), format = "latex")
```

@@

```{r}
#| label: tbl-Q-P-PREMISES_TYPE
#| echo: false
#| warning: false
#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of premise types"
monthly_counts_PREMISES_TYPE <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH,PREMISES_TYPE) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.Date(paste0(YEARMONTH, "-01")),
    action = ifelse(time >= as.Date("2024-01-01"),1,0))
poisson_model_PREMISES_TYPE <- glm(
  Crime_Count ~ time+PREMISES_TYPE +action,
  family = quasipoisson(),
  data = monthly_counts_PREMISES_TYPE
)
kable(summary(poisson_model_PREMISES_TYPE)$coefficients, format = "latex")
```

```{r}
#| label: tbl-Q-P-OFFENCE
#| echo: false
#| warning: false
#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of offence types"
monthly_counts_OFFENCE <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH,OFFENCE) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.Date(paste0(YEARMONTH, "-01")),
    action = ifelse(time >= as.Date("2024-01-01"),1,0))
poisson_model_OFFENCE <- glm(
  Crime_Count ~ time+OFFENCE +action,
  family = quasipoisson(),
  data = monthly_counts_OFFENCE
)
kable(summary(poisson_model_OFFENCE)$coefficients, format = "latex")
```

```{r}
#| label: tbl-Q-P-DIVISION
#| echo: false
#| warning: false
#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of divisions"
monthly_counts_DIVISION <- dt %>%
  mutate(
    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
  ) %>%
  group_by(YEARMONTH,DIVISION) %>%
  summarise(Crime_Count = n(), .groups = "drop") %>%
  mutate(
    time = as.Date(paste0(YEARMONTH, "-01"))  ,
    action = ifelse(time >= as.Date("2024-01-01"),1,0))
poisson_model_DIVISION <- glm(
  Crime_Count ~ time+DIVISION +action,
  family = quasipoisson(),
  data = monthly_counts_DIVISION
)
kable(summary(poisson_model_DIVISION)$coefficients, format = "latex")
```
